### 2014-03-19 Docs/flowchart/flowchart.tex
add the notation needed for the flowchart from the multilevel paper

reach\_state

enforce\_hse

@@

Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]$\\
\phantom{ }\hfill $\rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.

@@

Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$\\
\phantom{ }\hfill$\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$

@@

\item Compute $S^{\nph,\star}$.  First, compute $S^{n+1,\star}$ with
\begin{equation}
S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
\end{equation}
  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
  the equation of state.  Then, define
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} (S^{\nph,\star}),
\qquad
 S^{\nph.\star} = \frac{S^n + S^{n+1,\star}}{2},
\end{equation}
\item Compute $w_0^{\nph}$.  First, define
\begin{equation}
\overline{\Gamma_1^{\nph,\star}} = \frac{\overline{\Gamma_1^n} + \overline{\Gamma_1^{n+1,\star}}}{2}, 
\quad
\rho_0^{\nph,\star} = \frac{\rho_0^{n} + \rho_0^{n+1,\star}}{2},
\quad
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2},
\end{equation}
  with
\begin{equation}
 \overline{\Gamma_1^{n+1,\star}} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right].
\end{equation}
\begin{description}
\item For planar geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},p_0^{\nph,\star},\psi^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\item For spherical geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]$\\
\phantom{ }\hfill$\rightarrow [w_0^{\nph}]$.
\end{description}

>>>>

{\bf React State}$[\rho^{\inp},(\rho                                                                   h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp},                                                      
  p_0^{\inp}] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp},                                   
  T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
  evolves the species and enthalpy due to reactions through
  $\Delta t/2$ according to:
\begin{equation}
\frac{dX_k}{dt} = \omegadot_k(\rho,X_k,T) ; \qquad
\frac{dT}{dt}   = \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
\end{equation}
%This system is solved using the stiff ordinary differential equation                                     
%integration package, {\tt VODE} \citep{vode}.                                                            
Full details of the
solution procedure can be found in Paper III. We then define:
\begin{eqnarray}
(\rho\omegadot_k)^{\outp} &=& \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\dt/2}, \\
(\rho h)^{\outp} &=& (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\in\
p}.
\end{eqnarray}
where the enthalpy update includes external heat sources $(\rho\Hext)^{\inp}$.
As introduced in Paper IV, we update the temperature using $T^{\outp} =                                   
T(\rho^\outp,h^\outp,X_k^\outp)$ for planar geometry or $T^{\outp} =                                      
T(\rho^\outp,p_0^\inp,X_k^\outp)$ for spherical geometry.
Note that the density remains unchanged within {\bf React State}, i.e.,
$\rho^{\outp} = \rho^{\inp}$.

{\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp] \rightarrow
  [\rho_0^\outp, \rho_0^{\outp,\nph}]$ is the process by which we
  update the base state density through $\dt$ in time.  We keep the
  time-centered edge states, $\rho_0^{\outp,\nph}$,
  since they are used later in discretization of $\etarho$ for planar problems.
\begin{description}
\item[planar:] We discretize equation (\ref{eq:Base State Density}) to
  compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} \left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ].
\end{equation}
  We compute the time-centered edge states, $\rho_0^{\outp,\nph}$,
  by discretizing an expanded form of equation (\ref{eq:Base State Density}):
\begin{equation}
\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r},
\end{equation}
  where the right hand side is used as the force term.
\item[spherical:] The base state density update now includes the area factors in the 
  divergences:
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{1}{r_j^2} \frac{\dt}{\dr} \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right].
\end{equation}
  In order to compute the time-centered edge states, an additional geometric 
  term is added to the forcing, due to the spherical discretization of 
  (\ref{eq:Base State Density}):
\begin{equation}
\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r} - \frac{2 \rho_0 w_0}{r}.
\end{equation}
\end{description}

{\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp}]
  \rightarrow [p_0^{\outp}]$ has replaced {\bf Advect Base Pressure}
  from Paper III as the process by which we update the base state
  pressure.  Rather than discretizing the evolution equation for
  $p_0$, we enforce hydrostatic equilibrium directly, which is numerically simpler
  and analytically equivalent.  We first set 
  $p_{0,j=0}^{\outp} = p_{0,j=0}^{\inp}$ and then update $p_0^\outp$ using:
\begin{equation}
p_{0,j+1}^{\outp} = p_{0,j}^{\outp} + \Delta r g_{j+\myhalf}\frac{\left(\rho_{0,j+1}^{\inp}+\rho_{0,j}^{\inp}\right)}{2},
\end{equation}
  where $g=g(\rho_0^\inp)$.  As soon as $\rho_{0,j}^\inp < \rho_{\rm cutoff}$, we set 
  $p_{0,j+1}^\outp = p_{0,j}^\outp$ for all remaining values of $j$.  
  Then we compare $p_{0,j_{\rm max}}^\outp$ with $p_{0,j_{\rm max}}^\inp$ and offset
  every element in $p_0^\outp$ so that $p_{0,j_{\rm max}}^\outp = p_{0,j_{\rm max}}^\inp$.  
  We are effectively using the location where the $\rho_0^\inp$ drops below 
  $\rho_{\rm cutoff}$ as the starting point for integration.

{\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$
  is the process by which we update the base state enthalpy through $\dt$ in time.  
\begin{description}
\item[planar:] We discretize equation (\ref{eq:Base State Enthalpy}), neglecting reaction source terms,
  to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r} \left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} + \dt\psi_j^{\inp}.
\end{equation}
  We compute the time-centered edge states, $(\rho h)_0^{\nph}$, by discretizing
  an expanded form of equation (\ref{eq:Base State Enthalpy}):
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} + \psi.
\end{equation}
\item[spherical:]  The base state enthalpy update now includes the area factors 
  in the divergences:
\begin{eqnarray}
(\rho h)_{0,j}^{\outp} &=& (\rho h)_{0,j}^{\inp} \nonumber \\
&& - \frac{1}{r_j^2} \frac{\dt}{\dr} \left \{ \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} +\dt\psi^{\inp,\nph}.\nonumber\\
\end{eqnarray}
  In order to compute the time-centered edge states, an additional geometric 
  term is added to the forcing, due to the spherical discretization of 
  (\ref{eq:Base State Enthalpy}):
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} - \frac{2 (\rho h)_0 w_0}{r} + \psi.
\end{equation}
\end{description}

\subsubsection{Computing $w_0$}\label{Sec:Computing w0}
Here we describe the process by which we compute $w_0$.  The arguments 
are different for planar and spherical geometries.
\begin{description}

\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}
$[\Sbar^{\inp},\gammabar^{\inp}, p_0^{\inp},\psi^{\inp}]\rightarrow [w_0^{\outp}]$:

In Paper III, we showed that $\psi=\etarho g$ for planar geometries, and 
derived an alternate expression for equation (\ref{eq:w0 divergence}).
We compute $w_0$ using equation (35) in Paper III using the following 
discretization:
\begin{equation}
\frac{w_{0,j+\myhalf}^\outp-w_{0,j-\myhalf}^\outp}{\Delta r} = \left(\Sbar^{\inp} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}\right)_j,
\end{equation}
with $w_{0,-\myhalf}=0$.

\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}
  $[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},\etarho^{\inp}]
  \rightarrow[w_0^{\outp}]$:

We begin with equation (\ref{eq:w0 divergence}) written in spherical coordinates:
\begin{equation}
\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) = \beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ).
\end{equation}
We expand the spatial derivative and recall from Paper I that
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r},
\end{equation}
giving:
\begin{equation}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) = \Sbar - \frac{1}{\gammabar p_0} \underbrace{\left( \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)}_{\psi}.\label{eq:psi def}
\end{equation}
We solve this equation for $w_0$ as described in Appendix of the multilevel paper.

@@

Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]
 \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.

@@

Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$ \\
 ${}\hfill   \rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$

@@

\item Compute $S^{\nph,\star}$.  First, compute $S^{n+1,\star}$ with
\begin{equation}
S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
\end{equation}
  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
  the equation of state.  Then, define
\begin{equation}
\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} (S^{\nph,\star}),
\qquad
 S^{\nph.\star} = \frac{S^n + S^{n+1,\star}}{2},
\end{equation}
\item Compute $w_0^{\nph}$.  First, define
\begin{equation}
\overline{\Gamma_1^{\nph,\star}} = \frac{\overline{\Gamma_1^n} + \overline{\Gamma_1^{n+1,\star}}}{2}, 
\quad
\rho_0^{\nph,\star} = \frac{\rho_0^{n} + \rho_0^{n+1,\star}}{2},
\quad
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2},
\end{equation}
  with
\begin{equation}
 \overline{\Gamma_1^{n+1,\star}} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right].
\end{equation}
\begin{description}
\item For planar geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},p_0^{\nph,\star},\psi^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\item For spherical geometry, call\\
{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
\end{description}
